---
title: "Extreme Precipitation in the Netherlands: GEV Modelling & Extreme Value Copulas"
author: "Richard Marek (Condensed English version of a Master's thesis practical chapter)"
date: today
bibliography: references.bib
format:
  html:
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    code-fold: true
    code-tools: true
    theme: cosmo
    highlight-style: github
    embed-resources: true
    fig-width: 9
    fig-height: 5
    fig-align: center
    df-print: kable
execute:
  warning: false
  message: false
  cache: true
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(maps)
library(geosphere)
library(tidyr)
library(lubridate)
library(extRemes)
library(gridExtra)
library(copula)
library(ggrepel)
library(evd)
library(GGally)
library(reshape2)
library(stringr)
library(knitr)
library(kableExtra)
library(purrr)
```

---

## Why This Matters: Application to Natural Catastrophe Modelling {#motivation}

Before presenting the methodology, it is useful to briefly outline the relevance of extreme value theory and copula-based dependence modelling in **natural catastrophe (nat cat) modelling** within the insurance and reinsurance industry.

Extreme precipitation events are a key driver of flood losses in property insurance portfolios. Modern catastrophe models decompose risk into hazard, vulnerability, exposure, and financial loss components.

This analysis focuses on the hazard component by modelling both the marginal behaviour of extreme precipitation and its spatial dependence structure. The resulting framework can be used to generate stochastic event sets for portfolio-level risk assessment.

Nat cat models — developed by vendors such as Moody's (formerly RMS), Verisk
(formerly AIR Worldwide), and PERILS — decompose catastrophic risk into four components:

| Component | Description |
|-----------|-------------|
| **Hazard** | Probability of a damaging natural event at a given location and time |
| **Vulnerability** | Degree of loss as a function of event intensity |
| **Element at risk** | The exposed portfolio (insured values, locations, construction types) |
| **Risk** | Expected loss = Hazard × Vulnerability |

The **hazard module** is where extreme value theory comes in. GEV distributions and extreme
value copulas are the natural tools for modelling the frequency and severity of extreme
precipitation — and, critically, whether extremes at multiple locations occur *simultaneously*.

**Why simultaneous extremes matter for pricing:** A flood affecting a single city is
manageable. A flood affecting five cities at once can exhaust reinsurance capacity.
The key insight from this analysis — demonstrated numerically in the copula section — is that
even weak dependence between stations raises the probability of simultaneous exceedance
by **14× to 98× compared to the independence assumption**. Ignoring this leads to
systematic underestimation of aggregate Probable Maximum Loss (PML) and incorrect
Excess of Loss pricing.

### Directions for Further Development

The framework used here covers **marginal GEV fitting + bivariate extreme value copulas**,
corresponding to the first layer of a full nat cat hazard model. Natural extensions include:

- **Vine copulas** — decompose multivariate dependence into a cascade of bivariate copulas,
  one per pair (possibly conditional), allowing different copula families for each pair.
  This is far more flexible than a single 5D Gumbel copula with one parameter.
  See @Simpson2021 for vine copulas in an extreme value context.

- **Hierarchical (nested) Archimedean copulas** — stations are grouped by similarity;
  within-group dependence uses a copula with its own parameter, groups are then joined
  by a higher-level copula. The Gumbel copula is the only family that is simultaneously
  Archimedean and an extreme value copula, making it the natural choice here.

- **Copula averaging** — rather than selecting a single best copula, combine candidates
  as a weighted average (weights from AIC or BIC). The resulting model is not
  max-stable, but is a valid copula and is more robust under model uncertainty.

- **Spatial Bayesian hierarchical models** — GEV parameters (μ, σ, ξ) are modelled
  as spatially varying functions of location and covariates (e.g. elevation, distance
  from the coast). This allows *borrowing of information* between stations and produces
  full spatial return-level maps with uncertainty quantification. See @Bracken2021
  for a three-layer implementation: data layer (GEV + Gaussian copula), process layer
  (spatial regression per parameter), and spatial variability layer (radial basis functions).
  This type of output feeds directly into stochastic event set generation.

**From stochastic events to risk metrics:** Given a stochastic event set (typically
10,000+ synthetic years), each event is propagated through a vulnerability function to
produce insured losses per object. Aggregated across the portfolio, the loss distribution
supports calculation of:

- **AAL** (Average Annual Loss) — long-run expected annual loss; basis for rate-making
- **VaR** at 1-in-100 or 1-in-250 — regulatory capital and Excess of Loss attachment points
- **TVaR / Expected Shortfall** — mean loss in worst scenarios; a coherent risk measure

---

## Data

Daily precipitation totals were downloaded from the
[NOAA National Centers for Environmental Information (NCEI)](https://www.ncei.noaa.gov/products/land-based-station/global-historical-climatology-network-daily)
for five long-record stations in the Netherlands. The primary selection criterion was
record length and completeness, which is essential for reliable extreme value analysis.

```{r load-data}
data1 <- read.csv("4257416.csv", header = TRUE)

data1 <- data1 %>%
  mutate(
    NAME = gsub(",.*", "", NAME),
    NAME = gsub("\\d+", "", NAME),
    NAME = trimws(NAME),
    NAME = ifelse(NAME == "PUTTEN GLD", "PUTTEN", NAME),
    NAME = str_to_title(NAME),
    DATE = as.Date(DATE),
    PRCP = PRCP * 25.4   # inches → mm
  )

station_names <- unique(data1$NAME)

station_raw <- setNames(
  lapply(station_names, function(s) data1[data1$NAME == s, c("DATE", "PRCP")]),
  station_names
)

coords <- data1 %>%
  filter(NAME %in% station_names) %>%
  select(NAME, LATITUDE, LONGITUDE) %>%
  distinct()
```

### Station Map

```{r station-map, fig.cap="Location of the five meteorological stations in the Netherlands."}
nl_map <- map_data("world") %>% filter(region == "Netherlands")

ggplot() +
  geom_polygon(data = nl_map, aes(x = long, y = lat, group = group),
               fill = "lightyellow", color = "black") +
  geom_point(data = coords, aes(x = LONGITUDE, y = LATITUDE),
             color = "red", size = 3) +
  geom_text_repel(data = coords,
                  aes(x = LONGITUDE, y = LATITUDE, label = NAME),
                  size = 3.5, color = "black") +
  coord_fixed(1.3) +
  theme_minimal() +
  theme(axis.title = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), panel.grid = element_blank())
```

### Inter-station Distances

Stations at smaller geographic distances are expected to show stronger precipitation
dependence. Elevation ranges only from −3 to 20 m above sea level across all five
stations and is therefore not considered further.

```{r distances}
dist_matrix_km <- distm(coords[, c("LONGITUDE","LATITUDE")],
                        fun = distVincentyEllipsoid) / 1000

colnames(dist_matrix_km) <- rownames(dist_matrix_km) <- coords$NAME

dist_unique <- as.data.frame(dist_matrix_km) %>%
  mutate(STATION1 = coords$NAME) %>%
  pivot_longer(cols = -STATION1, names_to = "STATION2", values_to = "distance_km") %>%
  mutate(ST1 = pmin(STATION1, STATION2), ST2 = pmax(STATION1, STATION2)) %>%
  select(STATION1 = ST1, STATION2 = ST2, distance_km) %>%
  distinct() %>%
  filter(distance_km > 0) %>%
  arrange(distance_km) %>%
  mutate(distance_km = round(distance_km, 1))

kable(dist_unique,
      col.names = c("Station 1", "Station 2", "Distance (km)"),
      caption = "Pairwise distances between stations (Vincenty ellipsoid method).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped","hover"))
```

---

## Data Cleaning & Annual Maxima Extraction

```{r cleaning}
# EDA helper
compute_eda_row <- function(df, s, skip_yrs = character(0)) {
  n_na       <- sum(is.na(df$PRCP))
  full_dates <- seq(min(df$DATE), max(df$DATE), by = "day")
  if (length(skip_yrs) > 0)
    full_dates <- full_dates[!format(full_dates, "%Y") %in% skip_yrs]
  n_missing <- length(setdiff(full_dates, df$DATE))
  data.frame(Station       = s, Observations  = nrow(df),"Missing PRCP" = n_na,
  "Missing Days" = n_missing, "Start Date"= min(df$DATE),"End Date" = max(df$DATE),
             stringsAsFactors = FALSE)
}

# Identify problematic years (many missing values in 2021 and 2024)
SKIP_YEARS <- c("2021", "2024")

station_final <- setNames(
  lapply(station_names, function(s) {
    df         <- station_raw[[s]]
    first_year <- as.numeric(format(min(df$DATE), "%Y"))
    start_date <- as.Date(paste0(first_year, "-01-01"))
    if (min(df$DATE) > start_date)
      start_date <- as.Date(paste0(first_year + 1, "-01-01"))
    df <- df[!format(df$DATE, "%Y") %in% SKIP_YEARS, ]
    df[df$DATE >= start_date & df$DATE <= max(df$DATE, na.rm = TRUE), ]
  }),
  station_names
)

eda_final <- bind_rows(lapply(station_names, function(s)
  compute_eda_row(station_final[[s]], s, skip_yrs = SKIP_YEARS)))

kable(eda_final, caption = "Summary of cleaned daily data (years 2021 and 2024 excluded).") %>%
  kable_styling(bootstrap_options = c("striped","hover"))
```

Years 2021 and 2024 were excluded due to a large number of missing values across all
five stations. Where records did not begin on 1 January, the incomplete initial year
was discarded so that each block corresponds to a full calendar year.

```{r annual-maxima}
station_ymax <- setNames(
  lapply(station_names, function(s) {
    station_final[[s]] %>%
      mutate(YEAR = year(DATE)) %>%
      group_by(YEAR) %>%
      summarise(max_PRCP = max(PRCP, na.rm = TRUE), .groups = "drop")
  }),
  station_names
)
```

### Annual Maxima Time Series

```{r plot-maxima, fig.height=8, fig.cap="Annual maximum daily precipitation (mm) at each station."}
ymax_long <- bind_rows(lapply(station_names, function(s)
  station_ymax[[s]] %>% mutate(Station = s)))

ggplot(ymax_long, aes(x = YEAR, y = max_PRCP)) +
  geom_line(color = "steelblue") +
  geom_point(color = "red", size = 1.2) +
  facet_wrap(~ Station, ncol = 2) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(x = "Year", y = "Annual maximum precipitation (mm)") +
  theme_minimal()
```

---

## Assumption Checking for the Block Maxima Method

The block maxima method requires annual maxima to be independent and identically
distributed (i.i.d.). This cannot be verified directly; we therefore test necessary
conditions: **independence** (via ACF) and **stationarity** (via LOESS trend).

### Independence: Autocorrelation Functions

```{r acf-plots, fig.height=9, fig.cap="Sample autocorrelation functions of annual maximum precipitation."}
par(mfrow = c(3, 2), mar = c(4, 4, 3, 1))
for (s in station_names) {
  acf(station_ymax[[s]]$max_PRCP,
      main = s, xlab = "Lag (years)", ylab = "ACF")
}
```

For most stations, autocorrelation values remain within the 95% confidence bands.
Occasional exceedances (Groningen, Hoofddorp, Roermond) occur at higher lags without
a clear physical interpretation and are consistent with random chance given the number
of lags tested. No statistically significant autocorrelation is observed at low lags,
so the independence assumption is not rejected at the 5% level.

### Stationarity: LOESS Trend in Annual Maxima

```{r loess-maxima, fig.height=8, fig.cap="Annual maxima with LOESS trend (shaded 95% CI)."}
ggplot(ymax_long, aes(x = YEAR, y = max_PRCP)) +
  geom_point(color = "red", alpha = 0.6, size = 1.2) +
  geom_smooth(method = "loess", se = TRUE, color = "steelblue") +
  facet_wrap(~ Station, ncol = 2) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(x = "Year", y = "Annual maximum precipitation (mm)") +
  theme_minimal()
```

None of the five stations exhibits a strictly constant mean over time. The most
pronounced non-stationarity is at **Roermond** (~10 mm upward trend) and
**Hoofddorp** (strong recent increase). **Groningen** is closest to stationary behaviour.
**Putten** shows an inverted-U pattern, **Den Helder** a non-monotone curve.

A non-stationary GEV model with time-varying parameters would be more rigorous for these
stations (see @Katz for methods), but is beyond the scope of this analysis.
Results should be interpreted with this limitation in mind.

### LOESS Trend in Daily Data (All Observations)

```{r loess-daily, fig.height=8, fig.cap="LOESS trend fitted to all daily precipitation observations (individual points omitted due to volume)."}
daily_long <- bind_rows(lapply(station_names, function(s)
  station_final[[s]] %>% mutate(Station = s)))

ggplot(daily_long, aes(x = DATE, y = PRCP)) +
  geom_smooth(method = "loess", se = FALSE, color = "steelblue") +
  facet_wrap(~ Station, ncol = 2) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(x = "Year", y = "Daily precipitation (mm)") +
  theme_minimal()
```

Comparing daily and annual-maxima trends reveals that changes in overall precipitation
volumes do not always translate directly into changes in extreme behaviour.

---

## GEV Distribution Fitting

### Parameter Estimates

Annual maxima at each station are modelled with a Generalised Extreme Value (GEV)
distribution, $F(x;\mu,\sigma,\xi)$, estimated by maximum likelihood.

```{r gev-fit}
params_table <- map_dfr(station_names, function(s) {
  x   <- station_ymax[[s]]$max_PRCP
  fit <- fgev(x)
  ci  <- confint(fit)
  tibble(Station = s,
         mu = fit$est[1], sigma = fit$est[2], xi = fit$est[3],
         mu_low = ci[1,1], mu_high = ci[1,2],
         sigma_low = ci[2,1], sigma_high = ci[2,2],
         xi_low = ci[3,1], xi_high = ci[3,2])
})

params_table %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  kable(
    col.names = c("Station",
                  "μ", "σ", "ξ",
                  "low", "high", "low", "high", "low", "high"),
    caption = "GEV parameter estimates with 95% confidence intervals."
  ) %>%
  add_header_above(c(" " = 1,
                     "Estimate" = 3,
                     "μ CI" = 2,
                     "σ CI" = 2,
                     "ξ CI" = 2)) %>%
  kable_styling(bootstrap_options = c("striped","hover"))
```

The location parameter $\hat\mu$ is highest at Hoofddorp (29.68 mm). For Groningen,
Roermond, and Hoofddorp, the 95% CI for $\xi$ includes zero, making the Gumbel
distribution statistically acceptable. Den Helder and Putten show a mildly positive
$\hat\xi$ with CI excluding zero, indicating a slightly heavier upper tail.

### Diagnostic Plots

```{r gev-diagnostics, fig.height=10, fig.cap="GEV diagnostic plots for each station (histogram, KDE, ECDF, Q–Q, P–P)."}
plot_station_gev <- function(s, station_ymax, params_table) {
  x  <- station_ymax[[s]]$max_PRCP
  pars  <- params_table[params_table$Station == s, ]
  mu    <- pars$mu; sigma <- pars$sigma; xi <- pars$xi
  m     <- length(x)
  x_seq <- seq(min(x) * 0.9, max(x) * 1.15, length.out = 300)
  gev_df <- data.frame(x = x_seq,
                       pdf = dgev(x_seq, mu, sigma, xi),
                       cdf = pgev(x_seq, mu, sigma, xi))

  p1 <- ggplot(data.frame(x = x), aes(x = x)) +
    geom_histogram(aes(y = after_stat(density)), bins = 15,
                   fill = "steelblue", alpha = 0.5, color = "white") +
    geom_line(data = gev_df, aes(x = x, y = pdf), color = "red", linewidth = 1) +
    labs(x = "Annual max. precipitation", y = "Density") + theme_minimal()

  p2 <- ggplot(data.frame(x = x), aes(x = x)) +
    geom_density(fill = "steelblue", alpha = 0.4, color = "steelblue") +
    geom_line(data = gev_df, aes(x = x, y = pdf), color = "red", linewidth = 1) +
    labs(x = "Annual max. precipitation", y = "Density") + theme_minimal()

  ecdf_df <- data.frame(x = sort(x), ecdf = (1:m)/m)
  p3 <- ggplot(ecdf_df, aes(x = x, y = ecdf)) +
    geom_step(color = "steelblue", linewidth = 0.8) +
    geom_line(data = gev_df, aes(x = x, y = cdf), color = "red", linewidth = 1) +
    labs(x = "Annual max. precipitation", y = "CDF") + theme_minimal()

  qq_df <- data.frame(theoretical = qgev((1:m - 0.5)/m, mu, sigma, xi),
                      empirical   = sort(x))
  p4 <- ggplot(qq_df, aes(x = theoretical, y = empirical)) +
    geom_point(color = "steelblue", alpha = 0.7) +
    geom_abline(slope = 1, intercept = 0, color = "red", linewidth = 1) +
    labs(x = "Theoretical quantiles", y = "Empirical quantiles") + theme_minimal()

  pp_df <- data.frame(theoretical = pgev(sort(x), mu, sigma, xi),
                      empirical   = (1:m)/(m+1))
  p5 <- ggplot(pp_df, aes(x = theoretical, y = empirical)) +
    geom_point(color = "steelblue", alpha = 0.7) +
    geom_abline(slope = 1, intercept = 0, color = "red", linewidth = 1) +
    labs(x = "Theoretical probability", y = "Empirical probability") + theme_minimal()

  gridExtra::grid.arrange(
    p1, p2, p3, p4, p5,
    top = grid::textGrob(s, gp = grid::gpar(fontsize = 13, fontface = "bold")),
    ncol = 2)
}

for (s in station_names) plot_station_gev(s, station_ymax, params_table)
```

Overall, the GEV model fits well for Groningen, Roermond, and Hoofddorp. Den Helder
shows some underestimation of high quantiles (Q–Q plot, upper tail). Putten exhibits a
hint of bimodality, leading to larger Q–Q deviations in the middle range, and a clear
outlier in the upper tail where the theoretical model substantially overestimates.

### Return Level Plots

```{r return-levels, fig.height=9, fig.cap="Return level plots with 95% bootstrap confidence intervals (3rd and 97th percentile from 99 replications)."}
rl_ggplot <- function(s, station_ymax, params_table, xlim = c(1, 200)) {
  x   <- station_ymax[[s]]$max_PRCP
  fit <- fgev(x)
  pars <- params_table[params_table$Station == s, ]
  mu   <- pars$mu; sigma <- fit$param["scale"]; xi_s <- fit$param["shape"]

  ppx     <- ppoints(fit$tdata, a = 0)
  rps     <- c(1.001, 10^(seq(0, 3, len = 200))[-1])
  p_upper <- 1 / rps
  rlev    <- qgev(p_upper, loc = fit$loc, scale = sigma, shape = xi_s,
                  lower.tail = FALSE)

  df_pts  <- data.frame(rp  = -1/log(ppx),          obs  = sort(fit$tdata))
  df_line <- data.frame(rp  = -1/log(1 - p_upper),  rlev = rlev)

  # Bootstrap CI (3rd–97th percentile, 99 resamples)
  samp <- rgev(fit$n * 99, loc = fit$loc, scale = sigma, shape = xi_s)
  samp <- matrix(samp, fit$n, 99)
  samp <- apply(samp, 2, sort)
  samp <- apply(samp, 1, sort)
  env  <- t(samp[c(3, 97), ])
  df_ci <- data.frame(rp = -1/log(ppx), lo = env[,1], hi = env[,2])

  ggplot() +
    geom_errorbar(data = df_ci, aes(x = rp, ymin = lo, ymax = hi),
                  width = 0, color = "steelblue", alpha = 0.3) +
    geom_line(data = df_line, aes(x = rp, y = rlev)) +
    geom_point(data = df_pts,  aes(x = rp, y = obs), shape = 20) +
    scale_x_log10(breaks = c(1,2,5,10,20,50,100,200),
                  labels = c("1","2","5","10","20","50","100","200")) +
    coord_cartesian(xlim = xlim) +
    labs(title = s, x = "Return period (years)", y = "Return level (mm)") +
    theme_minimal()
}

plots_rl <- lapply(station_names, rl_ggplot,
                   station_ymax = station_ymax, params_table = params_table)
gridExtra::grid.arrange(grobs = plots_rl, ncol = 2)
```

Groningen and Hoofddorp show the tightest confidence bands and best model–data alignment.
Den Helder systematically underestimates high quantiles (empirical points above curve).
Putten shows very wide CIs for return periods > 50 years, reflecting high variability
in its input time series.

### GEV-based Risk Metrics

```{r gev-metrics}
THRESHOLD <- 80   # mm
K_YEAR    <- 100  # k-year return level

gev_metrics <- map_dfr(params_table$Station, function(s) {
  pars  <- params_table[params_table$Station == s, ]
  mu    <- pars$mu; sigma <- pars$sigma; xi <- pars$xi

  p_yr  <- 1 - pgev(THRESHOLD, mu, sigma, xi)  # annual exceedance prob.
  p_day <- 1 - pgev(THRESHOLD, mu, sigma, xi)^(1/365)  # daily exceedance prob.

  tibble(
    Station              = s,
    annual_exceedance    = round(p_yr, 4),
    daily_exceedance     = signif(p_day, 3),
    return_level_100yr   = round(qgev(1 - 1/K_YEAR, mu, sigma, xi), 2),
    return_period_80mm   = round(1/p_yr, 1)
  )
})

gev_metrics %>%
  kable(col.names = c("Station", "Annual P(X > 80 mm)",
                      "Daily P(X > 80 mm)",
                      "100-year return level (mm)",
                      "Return period for 80 mm (years)"),
        caption = "GEV-based extreme precipitation statistics (threshold = 80 mm).") %>%
  kable_styling(bootstrap_options = c("striped","hover"))
```

Putten has the highest annual exceedance probability (0.0078) with a return period of
~128 years for 80 mm, while Hoofddorp is the most conservative (return period ~1487 years).
The 100-year return level ranges from ~63 mm (Hoofddorp) to ~76 mm (Putten).

---

## Dependence Analysis

Before fitting copulas, we quantify pairwise dependence using two complementary approaches. Rank-based measures (Kendall's τ and Spearman's ρ) capture monotone dependence across the full distribution and are invariant to marginal transformations, so they can be applied to the original data directly. Tail dependence coefficients focus specifically on the probability that both variables are simultaneously extreme, which is the quantity most relevant for catastrophe risk. We also compare dependence estimated from annual maxima against daily data, as the two can differ substantially.

### Rank-based Correlation: Annual Maxima

```{r rank-corr}
# Common years across all stations
common_years <- Reduce(intersect, lapply(station_ymax, function(df) df$YEAR))
n_obs        <- length(common_years)
n_stations   <- length(station_names)

yearmax_common <- lapply(station_ymax, function(df) {
  df2 <- df[df$YEAR %in% common_years, ]
  df2[order(df2$YEAR), ]
})

X_matrix <- matrix(NA, nrow = n_obs, ncol = n_stations,
                   dimnames = list(common_years, station_names))
for (s in station_names)
  X_matrix[, s] <- yearmax_common[[s]]$max_PRCP
```

```{r kendall-tau, fig.cap="Kendall's τ between stations (annual maxima)."}
tau_matrix <- cor(X_matrix, method = "kendall")

tau_df <- as.data.frame(as.table(tau_matrix))
ggplot(tau_df, aes(Var1, Var2, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(Freq, 2)), size = 4) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                       midpoint = 0, name = expression(tau)) +
  coord_fixed() + theme_minimal() +
  labs(title = "Kendall's τ", x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r spearman-ymax, fig.cap="Spearman's ρ between stations (annual maxima)."}
spearman_ymax <- cor(X_matrix, method = "spearman")

spearman_df <- as.data.frame(as.table(spearman_ymax))
ggplot(spearman_df, aes(Var1, Var2, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(Freq, 2)), size = 4) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                       midpoint = 0, name = expression(rho)) +
  coord_fixed() + theme_minimal() +
  labs(title = "Spearman's ρ", x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

All pairwise dependences are positive but weak. Spearman's ρ ranges from 0.11 to 0.24,
Kendall's τ from 0.07 to 0.17. The strongest pair is Groningen–Putten (ρ = 0.24),
consistent with their shared exposure to similar weather systems.
Geography explains dependence only partially: Groningen–Putten (123 km) and
Groningen–Den Helder (127 km) are nearly equidistant, yet show different correlations.

### Spearman's ρ: Daily Data Comparison

```{r spearman-daily, fig.height=10, fig.cap="Spearman's ρ from daily data under three conditioning scenarios."}
common_dates <- Reduce(intersect, lapply(station_final, function(df) df$DATE))
common_dates <- as.Date(common_dates, origin = "1970-01-01")

D_matrix <- matrix(NA, nrow = length(common_dates), ncol = n_stations,
                   dimnames = list(as.character(common_dates), station_names))
for (s in station_names) {
  df <- station_final[[s]]
  df <- df[df$DATE %in% common_dates, ]
  D_matrix[, s] <- df[order(df$DATE), "PRCP"]
}

compute_spearman_rain <- function(D_matrix, mode = c("either","both")) {
  mode <- match.arg(mode)
  d    <- ncol(D_matrix); nms <- colnames(D_matrix)
  mat  <- matrix(NA, d, d, dimnames = list(nms, nms))
  for (i in 1:d) for (j in i:d) {
    x <- D_matrix[,i]; y <- D_matrix[,j]
    valid <- if (mode == "either") (x > 0 | y > 0) else (x > 0 & y > 0)
    val   <- cor(x[valid], y[valid], method = "spearman")
    mat[i,j] <- mat[j,i] <- val
  }
  mat
}

sp_all    <- cor(D_matrix, method = "spearman")
sp_either <- compute_spearman_rain(D_matrix, "either")
sp_both   <- compute_spearman_rain(D_matrix, "both")

plot_heatmap <- function(mat, title) {
  df <- as.data.frame(as.table(mat))
  ggplot(df, aes(Var1, Var2, fill = Freq)) +
    geom_tile(color = "white") +
    geom_text(aes(label = round(Freq, 2)), size = 3) +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                         midpoint = 0, name = expression(rho)) +
    coord_fixed() + theme_minimal() +
    labs(title = title, x = "", y = "") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

gridExtra::grid.arrange(
  plot_heatmap(sp_all,    "All days"),
  plot_heatmap(sp_either, "At least one station raining"),
  plot_heatmap(sp_both,   "Both stations raining"),
  ncol = 2
)
```

```{r rain-fraction}
rain_frac <- sapply(station_names, function(s)
  sum(station_final[[s]]$PRCP > 0) / nrow(station_final[[s]]))

kable(t(round(rain_frac, 3)),
      caption = "Fraction of days with positive precipitation at each station.") %>%
  kable_styling(full_width = FALSE)
```

Annual maxima show far weaker correlation (ρ ∈ [0.11, 0.24]) than daily data
(ρ ∈ [0.44, 0.71] for all days). The inflated all-day values are partly artefactual:
on 45–51% of days no rain falls at any station, creating many tied (0, 0) pairs
that artificially boost rank correlation. Conditioning on days when both stations rain
gives ρ ∈ [0.21, 0.59], closer to the true precipitation dependence.

This pattern — extremes showing weaker spatial dependence than everyday rain — is
well-documented and reflects that extreme precipitation is driven by different
meteorological mechanisms than routine events.

### Tail Dependence Coefficients

```{r tail-dep, fig.cap="Empirical upper (λᵤ) and lower (λₗ) tail dependence coefficients between all pairs."}
k_td <- round(sqrt(n_obs))  # ≈ √155 ≈ 12

lambda_lower <- function(X1, X2, k) {
  n <- length(X1); r <- rank(X1, ties.method = "average")
  s <- rank(X2, ties.method = "average")
  sum(r <= (n+1)/n*k & s <= (n+1)/n*k) / k
}
lambda_upper <- function(X1, X2, k) {
  n <- length(X1); r <- rank(X1, ties.method = "average")
  s <- rank(X2, ties.method = "average")
  thr <- n + 1 - (n+1)/n*k
  sum(r > thr & s > thr) / k
}

lam_l <- matrix(NA, n_stations, n_stations,
                dimnames = list(station_names, station_names))
lam_u <- matrix(NA, n_stations, n_stations,
                dimnames = list(station_names, station_names))

for (i in 1:n_stations) for (j in 1:n_stations) {
  if (i == j) { lam_l[i,j] <- 1; lam_u[i,j] <- 1
  } else {
    lam_l[i,j] <- lambda_lower(X_matrix[,i], X_matrix[,j], k_td)
    lam_u[i,j] <- lambda_upper(X_matrix[,i], X_matrix[,j], k_td)
  }
}

plot_tail <- function(mat, legend_label, title) {
  df <- reshape2::melt(mat, na.rm = TRUE)
  ggplot(df, aes(Var2, Var1, fill = value)) +
    geom_tile(color = "white", linewidth = 0.8) +
    geom_text(aes(label = round(value, 2)), size = 3.5, color = "black") +
    scale_fill_gradient(low = "#FFCCCC", high = "#CC0000",
                        limits = c(0, 1), name = legend_label) +
    coord_fixed() + labs(x = NULL, y = NULL, title = title) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 9))
}

gridExtra::grid.arrange(
  plot_tail(lam_u, expression(lambda[u]), expression("Upper tail dependence " ~ lambda[u])),
  plot_tail(lam_l, expression(lambda[l]), expression("Lower tail dependence " ~ lambda[l])),
  ncol = 2
)
```

Most pairs show $\hat\lambda_{u,n} = \hat\lambda_{l,n} = 0.08 = 1/12$, the minimum
non-zero value achievable with $k = 12$. Exceptions: Groningen–Putten, Roermond–Putten,
and Hoofddorp–Putten show $\hat\lambda_{u,n} = 0.17$–$0.25$. These estimates should be
treated as indicative only: with $n = 155$, $k \approx \sqrt{n} = 12$, and estimates
constrained to multiples of $1/12$, the resolution is very coarse.

As noted by @Serinaldi, standard estimators of $\hat\lambda_u$ can suffer from severe
bias and high variance. The apparent inconsistency between copula-selected dependence
strength and tail coefficient estimates is likely an artefact of estimator unreliability
rather than a true contradiction.

---

## Copula Analysis

By Sklar's theorem, any joint distribution can be decomposed into its marginal distributions and a copula that captures the dependence structure between them. This separation is particularly useful here: the marginal GEV distributions fitted in the previous section describe how extreme precipitation behaves at each station individually, while the copula describes how those extremes co-occur across stations. We focus on **extreme value copulas**, a subclass characterised by max-stability — the same property that makes GEV distributions the natural limit for block maxima. This theoretical consistency between the marginal and dependence models is what justifies the approach for extremes, as opposed to using general-purpose copulas such as the Gaussian or Clayton.

### Transformation to Uniform Marginals

Before fitting copulas, the marginal distribution of each station must be transformed to an approximate $U(0,1)$ distribution, in accordance with Sklar's theorem. Two approaches are implemented:

- **Parametric (IFM):**  

  $$\hat{u}_{i,j} = F_i(x_{i,j}; \hat{\phi}_i),$$
  
  where $F_i(\cdot  ; \hat{\phi}_i)$ is is the GEV distribution function with parameters    $\hat{\phi}_i= \left(\hat{\mu}, \ \hat{\sigma}, \ \hat{\xi}\right)$ estimated in the      previous section. This approach propagates uncertainty from the marginal fit into the     pseudo-observations.

- **Semi-parametric (MPL):** 

  $$\tilde{u}_{i,j} = \frac{n}{n+1} \hat{F}_i(x_{i,j}), \quad \hat{F}_i(x) = \frac{1}{n} \sum_{k=1}^n \mathbf{1}(x_{i,k} \le x). $$  
  Marginal distributions are left unspecified and observations are transformed via their    empirical ranks. The scaling factor $\frac{n}{n+1}$ (Hazen correction) ensures            pseudo-observations lie strictly inside $(0,1)$, avoiding numerical issues when           evaluating copula densities at the boundary — the standard empirical CDF with             denominator $n$ can attain the value 1 exactly.


```{r transform-U}
U_param <- matrix(NA, nrow = n_obs, ncol = n_stations,
                  dimnames = list(common_years, station_names))
for (s in station_names) {
  pars <- params_table[params_table$Station == s, ]
  U_param[, s] <- evd::pgev(yearmax_common[[s]]$max_PRCP,
                             loc = pars$mu, scale = pars$sigma, shape = pars$xi)
}

U_matrix <- matrix(NA, nrow = n_obs, ncol = n_stations,
                   dimnames = list(common_years, station_names))
for (s in station_names) {
  x <- yearmax_common[[s]]$max_PRCP
  m <- length(x)
  U_matrix[, s] <- rank(x, ties.method = "average") / (m + 1)
}
```

```{r hist-uniform, fig.cap="Histograms of transformed values for each station. Red line = U(0,1) reference density."}
plot_u_hist <- function(U, title) {
  as.data.frame(U) %>%
    pivot_longer(everything(), names_to = "station", values_to = "u") %>%
    ggplot(aes(x = u, y = after_stat(density))) +
    geom_histogram(binwidth = 0.1, fill = "steelblue", alpha = 0.6, color = "white") +
    geom_hline(yintercept = 1, color = "red", linewidth = 0.7) +
    scale_x_continuous(breaks = c(0, 0.5, 1)) +
    facet_wrap(~ station, nrow = 1) +
    labs(title = title, x = "Probability", y = "Density") +
    theme_minimal()
}

gridExtra::grid.arrange(
  plot_u_hist(U_param,  "Parametric transformation (IFM)"),
  plot_u_hist(U_matrix, "Semi-parametric transformation (MPL / Hazen)"),
  ncol = 1
)
```

The semi-parametric transformation produces histograms much closer to the uniform
reference, confirming that the fitted GEV distributions do not describe the empirical
distribution perfectly. All subsequent copula fitting uses the semi-parametric pseudo-observations.

```{r scatter-uniform, fig.cap="Pairwise scatter plots of semi-parametric pseudo-observations. Weak but non-zero dependence is visible."}
ggpairs(as.data.frame(U_matrix),
        lower = list(continuous = wrap("points", alpha = 0.3, size = 0.7, color = "steelblue")),
        diag  = list(continuous = wrap("densityDiag", fill = "steelblue", alpha = 0.4)),
        upper = list(continuous = wrap("points", alpha = 0.3, size = 0.7, color = "steelblue"))) +
  scale_x_continuous(breaks = c(0, 1)) +
  scale_y_continuous(breaks = c(0, 1)) +
  theme_minimal()
```

### Bivariate Extreme Value Copula Selection

For each of the $\binom{5}{2} = 10$ station pairs, five extreme value copula families
are fitted by maximum pseudo-likelihood (MPL) and ranked by AIC:
Gumbel, Galambos, Hüsler–Reiss, t-EV, and Tawn (symmetric implementation).

Note: The Joe extreme value copula is not available in the `copula` package (the Joe
Archimedean copula is a different family); it is therefore excluded.

```{r copula-functions}
draw_3d <- function(cop, ttl = "", n = 30) {
  u_seq <- seq(1e-4, 1 - 1e-4, length.out = n)
  grid  <- as.matrix(expand.grid(u1 = u_seq, u2 = u_seq))
  z_mat <- matrix(log(dCopula(grid, cop)), nrow = n)
  par(pty = "m")
  persp(u_seq, u_seq, z_mat, theta = 40, phi = 25, expand = 0.6,
        col = "steelblue", shade = 0.4, border = NA,
        xlab = "u", ylab = "v", zlab = "log c(u,v)", main = ttl,
        ticktype = "detailed")
}

draw_contour <- function(cop, ttl = "", n = 60) {
  u_seq <- seq(0.01, 0.99, length.out = n)
  grid  <- expand.grid(u1 = u_seq, u2 = u_seq)
  z_mat <- matrix(log(dCopula(as.matrix(grid), cop)), nrow = n)
  pal   <- hcl.colors(64, "YlOrRd", rev = TRUE)
  par(pty = "s")
  image(u_seq, u_seq, z_mat, col = pal, xlim = c(0,1), ylim = c(0,1),
        xlab = "u", ylab = "v", main = ttl)
  clip(0, 1, 0, 1)
  contour(u_seq, u_seq, z_mat, add = TRUE, nlevels = 10, col = "grey20", lwd = 0.6)
  box()
}

draw_pickands <- function(cop, ttl = "", n = 200) {
  s_seq  <- seq(0.001, 0.999, length.out = n)
  A_vals <- A(cop, s_seq)
  lower  <- pmax(s_seq, 1 - s_seq)
  plot(s_seq, A_vals, type = "n", xlim = c(0,1), ylim = c(0.25, 1.25),
       xlab = "s", ylab = "A(s)", main = ttl, asp = 1)
  polygon(c(s_seq, rev(s_seq)), c(rep(1, length(s_seq)), rev(lower)),
          col = "grey88", border = NA)
  lines(s_seq, lower, col = "tomato",      lwd = 1.5, lty = 2)
  abline(h = 1,       col = "forestgreen", lwd = 1.5, lty = 3)
  lines(s_seq, A_vals, col = "steelblue",  lwd = 2)
  legend("bottomleft", legend = c("A(s)", "max{s,1-s}", "1"),
         col = c("steelblue","tomato","forestgreen"),
         lwd = c(2,1.5,1.5), lty = c(1,2,3), bty = "n", cex = 0.75)
  box()
}

select_best_copula <- function(u1, u2) {
  umat <- cbind(u1, u2)
  candidates <- list(
    list(nm = "Gumbel",       cop0 = gumbelCopula(),      k = 1),
    list(nm = "Galambos",     cop0 = galambosCopula(),    k = 1),
    list(nm = "Husler-Reiss", cop0 = huslerReissCopula(), k = 1),
    list(nm = "tEV",          cop0 = tevCopula(),          k = 2),
    list(nm = "Tawn",         cop0 = tawnCopula(),         k = 2)
  )
  results <- data.frame(copula = character(), AIC = numeric(),
                        loglik = numeric(), k = integer(),
                        stringsAsFactors = FALSE)
  fitted  <- list()
  for (cand in candidates) {
    fit <- tryCatch(fitCopula(cand$cop0, umat, method = "mpl",
                              optim.control = list(maxit = 500)),
                   error = function(e) NULL)
    if (is.null(fit)) next
    ll <- tryCatch(logLik(fit)[1], error = function(e) NA)
    if (is.na(ll) || !is.finite(ll)) next
    aic <- 2 * cand$k - 2 * ll
    results <- rbind(results, data.frame(copula = cand$nm, AIC = aic,
                                         loglik = ll, k = cand$k))
    fitted[[cand$nm]] <- fit@copula
  }
  if (nrow(results) == 0) stop("No copula converged!")
  best_nm <- results$copula[which.min(results$AIC)]
  list(results = results, best = best_nm, fitted = fitted)
}
```

```{r fit-copulas, cache=TRUE}
pairs_list    <- combn(station_names, 2, simplify = FALSE)
selection_log <- data.frame()
fitted_pairs  <- list()

for (pair in pairs_list) {
  s1 <- pair[1]; s2 <- pair[2]
  sel <- select_best_copula(U_matrix[, s1], U_matrix[, s2])
  selection_log <- rbind(selection_log,
                         data.frame(Station1 = s1, Station2 = s2,
                                    Best_copula = sel$best,
                                    AIC = round(min(sel$results$AIC), 2)))
  fitted_pairs[[paste(s1, s2, sep = "_")]] <- sel$fitted[[sel$best]]
}
```

### Summary of Selected Copulas

```{r copula-summary-table}
selection_log %>%
  kable(col.names = c("Station 1", "Station 2", "Selected copula", "AIC"),
        caption = "Best-fitting extreme value copula per station pair (minimum AIC).") %>%
  kable_styling(bootstrap_options = c("striped","hover"))
```

Hüsler–Reiss was selected for 6 of 10 pairs, t-EV for 3, and Gumbel for 1
(Groningen–Den Helder). Geographically close pairs — particularly Hoofddorp–Putten
(63 km, t-EV) and Den Helder–Hoofddorp (69 km, t-EV) — show the strongest dependence,
captured by t-EV which allows for tail dependence. More distant pairs are described by
Hüsler–Reiss with weaker dependence. All Pickands functions lie in [0.88, 1.00],
confirming weak to moderate tail dependence throughout.

### Copula Diagnostic Plots (All Pairs)

```{r copula-diagnostics, fig.height=5, fig.cap="3D log-density, contour plot, and Pickands function A(s) for each station pair."}
for (pair in pairs_list) {
  s1  <- pair[1]; s2 <- pair[2]
  key <- paste(s1, s2, sep = "_")
  cop <- fitted_pairs[[key]]
  best_nm <- selection_log$Best_copula[selection_log$Station1 == s1 &
                                       selection_log$Station2 == s2]

  par(mfrow = c(1, 3), mar = c(4, 4, 3.5, 1), oma = c(0, 0, 3, 0))
  draw_3d(cop,      "3D log-density log c(u,v)")
  draw_contour(cop, "Contour plot log c(u,v)")
  draw_pickands(cop, "Pickands function A(s)")
  mtext(paste0(best_nm, "  [", s1, " × ", s2, "]  (selected by AIC)"),
        outer = TRUE, cex = 1.1, font = 2)
}
```

---

## Risk Metrics from Bivariate Copulas

We estimate the probability of simultaneously exceeding 80 mm at both stations
in a given year, and the corresponding joint return period. These cannot be
computed analytically for all families, so Monte Carlo simulation is used
($10^6$ draws per pair, seeds fixed for reproducibility).

```{r copula-probs, cache=TRUE}
u_vals  <- sapply(station_names, function(s) {
  pars <- params_table[params_table$Station == s, ]
  pgev(THRESHOLD, pars$mu, pars$sigma, pars$xi)
})
p_exceed <- 1 - u_vals
t_d1     <- 1 / p_exceed

p_simulation <- function(cop, u_vals, n_sim = 1e6, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  d   <- length(u_vals)
  sim <- rCopula(n_sim, cop)
  mean(rowSums(sim > matrix(u_vals, nrow = n_sim, ncol = d, byrow = TRUE)) == d)
}

pair_p <- matrix(NA, n_stations, n_stations,
                 dimnames = list(station_names, station_names))
pair_T <- matrix(NA, n_stations, n_stations,
                 dimnames = list(station_names, station_names))
diag(pair_p) <- p_exceed
diag(pair_T) <- t_d1

for (i in 1:(n_stations - 1)) for (j in (i+1):n_stations) {
  s1 <- station_names[i]; s2 <- station_names[j]
  cop_ij <- fitted_pairs[[paste(s1, s2, sep = "_")]]
  p_ij   <- p_simulation(cop_ij, c(u_vals[i], u_vals[j]),
                         n_sim = 1e6, seed = 123 + i*10 + j)
  pair_p[i,j] <- pair_p[j,i] <- p_ij
  pair_T[i,j] <- pair_T[j,i] <- 1 / p_ij
}
```

```{r bivariate-prob-table}
round(pair_p, 5) %>%
  kable(caption = "Joint annual probability of simultaneously exceeding 80 mm.
        Diagonal = marginal exceedance probabilities.") %>%
  kable_styling(bootstrap_options = c("striped","hover"))
```

```{r bivariate-rt-table}
round(pair_T, 1) %>%
  kable(caption = "Joint return period for simultaneous exceedance of 80 mm (years).
        Diagonal = marginal return periods.") %>%
  kable_styling(bootstrap_options = c("striped","hover"))
```

```{r ratio-table}
ratio_pairs <- matrix(NA, n_stations, n_stations,
                      dimnames = list(station_names, station_names))
for (i in 1:(n_stations-1)) for (j in (i+1):n_stations) {
  ratio_pairs[i,j] <- pair_p[i,j] / (p_exceed[i] * p_exceed[j])
  ratio_pairs[j,i] <- ratio_pairs[i,j]
}

round(ratio_pairs, 1) %>%
  kable(caption = "Ratio of copula-implied joint probability to independence assumption.
        Values >> 1 indicate that ignoring dependence severely underestimates risk.") %>%
  kable_styling(bootstrap_options = c("striped","hover"))
```

**Key finding for cat modelling:** Although the fitted copulas exhibit only weak tail dependence — with Pickands functions lying close to 1 throughout — the copula-implied joint exceedance probability is **14× to 98× higher** than under the independence assumption. The Hoofddorp–Groningen pair shows the largest amplification (~98×), despite having one of the lower Spearman correlations (ρ = 0.14). This illustrates a general property of extreme value copulas: even marginal dependence in the tail can have an outsized effect on joint exceedance probabilities, because both marginal probabilities are already very small and any positive association compounds multiplicatively.
In a cat modelling context, this matters at every stage of the risk chain. An excess of loss reinsurance structure priced under independence would systematically understate the probability of simultaneous losses across locations, leading to insufficient attachment points and underpriced aggregate covers. PML estimates at the 1-in-200 level — used for regulatory capital under Solvency II — would similarly be understated.

---

## 5-Dimensional Gumbel Copula

The `copula` package supports only the Gumbel copula for dimensions > 2 among extreme
value families. A 5D Gumbel copula is fitted to capture the joint behaviour of all
five stations simultaneously.

```{r fit-5d, cache=TRUE}
fit_5d <- fitCopula(gumbelCopula(dim = 5), U_matrix, method = "mpl",
                    optim.control = list(maxit = 1000))
cop5d <- fit_5d@copula

cat(sprintf("θ = %.4f,  Kendall's τ = %.4f,  loglik = %.2f,  AIC = %.2f\n",
            coef(fit_5d), tau(cop5d), logLik(fit_5d)[1],
            2*1 - 2*logLik(fit_5d)[1]))

P_5d       <- p_simulation(cop5d, u_vals, seed = 123)
P_indep_5d <- prod(p_exceed)
ratio_5d   <- P_5d / P_indep_5d

fiveD_summary <- tibble(
  Description = c("Joint P(all 5 > 80 mm) — copula",
                  "Joint P(all 5 > 80 mm) — independence",
                  "Joint return period (years)",
                  "Ratio copula / independence"),
  Value = c(P_5d, P_indep_5d, 1/P_5d, ratio_5d)
)

kable(fiveD_summary, digits = c(0, 10),
      caption = "5-dimensional Gumbel copula.") %>%
  kable_styling(bootstrap_options = c("striped","hover"))
```

The estimated $\hat\theta = 1.12$ is close to 1 (independence), yet the copula-implied
joint probability is **~239 million times higher** than under independence.
This demonstrates how, in high dimensions, even negligible dependence dramatically
amplifies simultaneous extreme event probabilities — a critical consideration when
pricing multi-location aggregate covers.

**Caution:** Although the fitted 5D Gumbel copula indicates only very weak dependence (θ close to 1), the implied joint exceedance probability is several orders of magnitude higher than under independence. This highlights a key feature of high-dimensional extreme value modelling: even weak positive dependence can substantially amplify joint tail risk. The result should therefore be interpreted as an **illustration of sensitivity to dependence assumptions**, rather than as a fully reliable estimate due to the restrictive copula structure.

---

## Conclusions

The analysis proceeds from individual station extremes to multivariate risk, following the structure of a catastrophe model hazard component. The main findings are summarised below, with emphasis on their practical implications for aggregate risk estimation.

1. **Marginal extremes:** Annual maxima at all five stations are well described by GEV
   distributions. Den Helder and Putten show heavier upper tails (ξ > 0 with CI excluding
   zero). Non-stationarity in the mean is present at all stations, most markedly at
   Roermond and Hoofddorp, which should be noted when interpreting return periods.

2. **Dependence structure:** All pairwise dependencies are positive but weak
   (Spearman's ρ ∈ [0.11, 0.24]). Proximity matters but is not the only driver.
   Extremes show weaker spatial correlation than daily totals, consistent with
   extreme precipitation being driven by localised meteorological systems.

3. **Copula selection:** Hüsler–Reiss dominates (6/10 pairs), t-EV appears for the
   three geographically closest pairs. All Pickands functions lie near 1, indicating
   weak but non-zero upper tail dependence.

4. **Risk amplification:** Despite weak dependence, the copula-implied joint exceedance
   probability is 14–98× higher than under independence for station pairs, and ~239
   million× higher in the 5D case. **These findings indicate that ignoring spatial
   dependence in extreme precipitation modelling may lead to systematic underestimation 
   of aggregate tail risk in excess-of-loss pricing structures.**

Taken together, these results suggest that a naive independence assumption — which remains common in simplified pricing models — can understate joint exceedance probabilities by one to two orders of magnitude even when pairwise correlations appear negligible. The vine copula and spatial hierarchical extensions outlined in the introduction would allow this framework to scale to larger station networks and produce the spatial return-level maps required for full stochastic event set generation.

---

## References

::: {#refs}
:::

```{r session-info, include=FALSE}
sessionInfo()
```
